---
title: "ht5"
author: "Dmitriy Tsimokha"
date: "07.12.20"
output:
  rmdformats::downcute:
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: true
---

<style type="text/css">
  body{
  font-size: 14pt;
}
</style>

**Turn the dark theme if you are reading it at night! (button with sun at the left)**

```{r data_loading, message=FALSE, warning=FALSE}
library(foreign); library(knitr); library(ggplot2); library(kableExtra)
library(jtools); library(memisc); library(sjPlot); library(ggstance)
library(broom.mixed); library(stargazer); library(ggthemr)
df <- read.spss("dataht5.sav", use.value.labels = T, to.data.frame = T)
df <- na.omit(df[c("happy", "gndr", "inwyys", "yrbrn", "sclmeet")])
```

# 1. Visualizing distributions
```{r settings_viz, include=FALSE}
#devtools::install_github('Mikata-Project/ggthemr')
# https://github.com/Mikata-Project/ggthemr
# https://github.com/ricardo-bion/ggtech
# https://github.com/hrbrmstr/hrbrthemes
ggthemr("flat dark", type = "outer")
# For transparent plots:
# theme_update(panel.background = element_rect(fill = "transparent",
#                                              colour = NA),
#              plot.background = element_rect(fill = "transparent",
#                                             colour = NA))
#opts_chunk$set(dev.args=list(bg="transparent"))
```

[Really nice graphing guide!](https://cedricscherer.netlify.app/2019/08/05/a-ggplot2-tutorial-for-beautiful-plotting-in-r/)

[Even nicer amount of themes for plotting!](https://rfortherestofus.com/2019/08/themes-to-improve-your-ggplot-figures/)

*I just can't stop choosing between different themes, help*

## Happy (dependent variable)
```{r happy_viz}
g <- ggplot(df, aes(x = happy))
g + geom_bar(stat = 'count') + 
  labs(x = "Level of happiness", y = "Count")
```

**Level of happiness** looks like skewed to the right distribution, but can be considered as quite close to normal distribution.

## Gender
```{r gender_viz}
g <- ggplot(df, aes(x = gndr))
g + geom_bar(stat = 'count') + 
  labs(x = "Gender", y = "Count")
```

There are **more female** respondents than male ones - nearly on one fourth larger sample.

## Age
```{r age_viz}
df$age <- as.numeric(levels(df$inwyys))[df$inwyys] - as.numeric(levels(df$yrbrn))[df$yrbrn]
ggplot(df, aes(x = age)) + 
  geom_density(fill = "salmon", stat = 'density') + 
  stat_function(fun = dnorm, args = list(mean = mean(df$age), sd = sd(df$age))) +
  labs(x = "Age", y = "Density")
```

Distrubution of **age** have two peakes at ~30 and at ~60 but still can be considered as close to equal (if we really want to).

## Meeting with friends and colleagues frequency
```{r meeting_viz}
ggplot(df, aes(x = sclmeet)) + 
  geom_bar(stat = 'count') + 
  labs(x = "Frequency of meeting", y = "Count")
```

Too can be considered as normal distribution (*if look from far distance*).


# 2. Simple model with control variables (age & gender)

[Great visualization tool for regression!](https://cran.r-project.org/web/packages/jtools/vignettes/summ.html)

```{r message=FALSE, warning=FALSE}
model1 <- lm(data = df, as.numeric(happy) ~ age + gndr)
tab_model(model1)
```

And **gender** is not significant! It can be seen not only by p-value larger than 0.05, but also that confident interval includes **0**.

Even though **age** is significant it still causes very small effect at first glance - but if we consider that age don't start from 0 and difference can be 50 or more years difference in effect can be really large (per -0.02 for each year).

*Looks like the older we get the less happier we feel :(*


# 3. More complex model with meeting frequency

```{r message=FALSE, warning=FALSE}
model2 <- lm(data = df, as.numeric(happy) ~ age + gndr + sclmeet)
# https://strengejacke.github.io/sjPlot/articles/tab_model_estimates.html
tab_model(model2)
```

The second model is more complex and include **frequency of meeting** with friends and colleagues:
- **gender** again is not significant! 
- **frequency of meeting** is significant and have **strong positive** effect on level of happiness! 

```{r results="asis", message=FALSE, warning=FALSE}
stargazer(anova(model1, model2), type = "html", title="ANoVA", intercept.bottom = FALSE,
          single.row = TRUE, align=TRUE, ci=TRUE, ci.level=0.95, no.space=TRUE)
```

The **model2** with **frequency of meeting** is significantly better than *model1*: lower RSS shows that.


```{r message=FALSE, warning=FALSE}
tab_model(model1, model2)
```

Still, each model describes only up to 0.04 variance max at the best case and that's too bad! 

# 4. Interpretation

## Assumptions
```{r}
plot_model(model2, type = "diag")[[1]]
```

We don't have any multicollinearity!

```{r message=FALSE, warning=FALSE}
plot_model(model2, type = "diag")[[2]]
```

And even residuals distributed normally and there are no major outliers!

```{r}
plot_model(model2, type = "diag")[[3]]
```

Here again - residuals at the best shape and very nicely fits under the normal distribution line.

```{r message=FALSE, warning=FALSE}
plot_model(model2, type = "diag")[[4]]
```

And here the baddest picture - homoscedasticity. Looks as bad as \[R^2\] of our models.

# 5. Visualization of the better model

*shine bright!*

```{r message=FALSE, warning=FALSE}
plot_summs(model2, plot.distributions = TRUE, inner_ci_level = .9)
```

Again, here can be seen coefficients with confidence intervals of the best (second) model:
- **age** is nearly on the line
- **gender** crosses the line with it's confidence interval
- and for different levels of **meeting frequency** everything great!

## Effects

```{r}
plot_model(model2, type = "pred")[[1]]
```

**Age** has significant negative effect and with greater age level of happiness expected to lower.

```{r}
plot_model(model2, type = "pred")[[2]]
```

**Gender** is not significant so that picture can be scipped.

```{r}
plot_model(model2, type = "pred")[[3]]
```

And **level of happiness** significantly affected by **meeting frequency** with positive direction!

# Summary

It does not matter what gender are you - just spend time with your friends more while you are young and **be happy**!
