---
title: "ht6"
author: "Dmitriy Tsimokha"
date: "13.12.20"
link: (click)[http://mglotov.me/da2020/basics/ht6.html]
output:
  rmdformats::downcute:
    code_folding: show
    self_contained: true
    thumbnails: false
    lightbox: true
---

<div class="snowflakes" aria-hidden="true">
  <div class="snowflake">
  ❅
  </div>
  <div class="snowflake">
  ❅
  </div>
  <div class="snowflake">
  ❆
  </div>
  <div class="snowflake">
  ❄
  </div>
  <div class="snowflake">
  ❅
  </div>
  <div class="snowflake">
  ❆
  </div>
  <div class="snowflake">
  ❄
  </div>
  <div class="snowflake">
  ❅
  </div>
  <div class="snowflake">
  ❆
  </div>
  <div class="snowflake">
  ❄
  </div>
</div>


```{css echo = FALSE}
body {
  font-size: 14pt;
}
/*Source: https://codepen.io/codeconvey/pen/xRzQay*/
/* customizable snowflake styling */
.snowflake {
  color: lightblue;
  font-size: 5em;
  font-family: Arial;
  text-shadow: 0 0 1px #000;
}
@-webkit-keyframes snowflakes-fall{0%{top:-10%}100%{top:100%}}
@-webkit-keyframes snowflakes-shake{
    0%{-webkit-transform:translateX(0px);transform:translateX(0px)}
   50%{-webkit-transform:translateX(80px);transform:translateX(80px)}
  100%{-webkit-transform:translateX(0px);transform:translateX(0px)}}
@keyframes snowflakes-fall{
    0%{top:-10%}
  100%{top:100%}}
@keyframes snowflakes-shake{
    0%{transform:translateX(0px)}
   50%{transform:translateX(80px)}
  100%{transform:translateX(0px)}}
.snowflake{
  position:fixed;
  top:-10%;
  z-index:9999;
  -webkit-user-select:none;
  -moz-user-select:none;
  -ms-user-select:none;
  user-select:none;
  cursor:default;
  -webkit-animation-name:snowflakes-fall,snowflakes-shake;
  -webkit-animation-duration:10s,3s;
  -webkit-animation-timing-function:linear,ease-in-out;
  -webkit-animation-iteration-count:infinite,infinite;
  -webkit-animation-play-state:running,running;
  animation-name:snowflakes-fall,snowflakes-shake;
  animation-duration:10s,3s;
  animation-timing-function:linear,ease-in-out;
  animation-iteration-count:infinite,infinite;
  animation-play-state:running,running}
.snowflake:nth-of-type(0){left: 1%;-webkit-animation-delay:  0s,  0s;animation-delay:  0s,  0s}
.snowflake:nth-of-type(1){left:10%;-webkit-animation-delay:  1s,  1s;animation-delay:  1s,  1s}
.snowflake:nth-of-type(2){left:20%;-webkit-animation-delay:  6s,0.5s;animation-delay:  6s,0.5s}
.snowflake:nth-of-type(3){left:30%;-webkit-animation-delay:  4s,  2s;animation-delay:  4s,  2s}
.snowflake:nth-of-type(4){left:40%;-webkit-animation-delay:  2s,  2s;animation-delay:  2s,  2s}
.snowflake:nth-of-type(5){left:50%;-webkit-animation-delay:  8s,  3s;animation-delay:  8s,  3s}
.snowflake:nth-of-type(6){left:60%;-webkit-animation-delay:  6s,  2s;animation-delay:  6s,  2s}
.snowflake:nth-of-type(7){left:70%;-webkit-animation-delay:2.5s,  1s;animation-delay:2.5s,  1s}
.snowflake:nth-of-type(8){left:80%;-webkit-animation-delay:  1s,  0s;animation-delay:  1s,  0s}
.snowflake:nth-of-type(9){left:90%;-webkit-animation-delay:  3s,1.5s;animation-delay:  3s,1.5s}
```


I've chosen **Switzerland** on **9th** round - just because someday I want to live there (*among mountains and forests!*)

```{r data_loading, message=FALSE, warning=FALSE}
library(ggthemr); library(foreign); library(ggplot2); library(stringr); 
library(sjPlot); library(stargazer); library(jtools)
ggthemr("lilac", type = "outer")
df <- read.spss("ESS9CH.sav", use.value.labels = T, to.data.frame = T)
df <- na.omit(df[c("plnftr", "trstlgl", "agea", "rshpsts", "bthcld", "gndr")])
df <- na.omit(df)
```


# Variables & model

**My hypothesis**: person's attitude towards *planning for future or taking each day as it comes* relates on how stable are situation in country with legal system which can be evaluates through person's *trust in the legal system* attitude - **more trust person have to legal system, more she/he perceive situation in country as stable therefore more planning for future**.

Such relation assumed to have different effect strength with people of different age and on people with/without partners and children: (*I understand that hypotheses for control variables are not used usially, but why not*)   
- **age**: simple case - more years, more planning; complex case - more years (till 50s), more planning & more years (after 50s), less planning    
- **partner**: if there are any partner now, assumed that person plan more    
- **children**: if there are any child(ren), assumed that person plan more    


# Variables

- **Dependent**: plnftr: Plan for future or take each day as it comes   
- **Predictor**: trstlgl: Trust in the legal system   
- *Control*: agea: Age of respondent, calculated    
- *Control*: rshpsts: Relationship with husband/wife/partner currently living with    
- *Control*: bthcld: Ever given birth to/ fathered a child    


## Variables distributions: plnftr

```{r}
ggplot(df, aes(x = plnftr)) + geom_bar(stat = "count") +
  labs(x = "Plan for future or take each day as it comes", y = "Count") + 
  scale_x_discrete(labels = function(x) str_wrap(x, width = 15))
```

So dependent variable distributed not really equally but nevertheless - I will analyse it.

## Variables distributions: trstlgl

```{r}
ggplot(df, aes(x = trstlgl)) + geom_bar(stat = "count") +
  labs(x = "Trust in the legal system", y = "Count")
```

There are strong skew to the right on the trust in the legal system for Switzerland - *that's good for them* (**but not good for the model**)!

## Variables distributions: agea

```{r}
ggplot(df, aes(x = as.numeric(agea))) + 
  geom_density(fill = "salmon", stat = 'density') + 
  stat_function(fun = dnorm, args = list(mean = mean(as.numeric(df$agea)), sd = sd(as.numeric(df$agea)))) +
  labs(x = "Age", y = "Density")
```

Age distributed normally as always in ESS (*good job*).

## Variables distributions: rshpsts

```{r}
ggplot(df, aes(x = rshpsts)) + geom_bar(stat = "count") +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 15)) +
  labs(x = "Relationship with husband/wife/partner currently living with", y = "Count")
```

And here will be a lot of problems with such distribution. So let's take **gender as other control** variable instead this one.

## Variables distributions: gndr

```{r}
ggplot(df, aes(x = gndr)) + geom_bar(stat = "count") +
  labs(x = "Gender", y = "Count")
```

Very nice (almost uniform) distribution.

## Variables distributions: bthcld

```{r}
ggplot(df, aes(x = bthcld)) + geom_bar(stat = "count") +
  labs(x = "Ever given birth to/ fathered a child", y = "Count")
```

Not really nice distribution but okay.


# Models

*Remember! I don't revert the scale of planning for future - zero is more planning attitude!*

## Comparison

```{r}
model1 <- lm(data = df, as.numeric(plnftr) ~ as.numeric(agea) + gndr + bthcld)
model2 <- lm(data = df, as.numeric(plnftr) ~ as.numeric(trstlgl))
# Main model
model <- lm(data = df, as.numeric(plnftr) ~ as.numeric(trstlgl) + as.numeric(agea) + gndr + bthcld)
tab_model(model, model1, model2, title = "Models comparison", dv.labels = c("predictor & control", "Only control", "Only predictor"))
```

As can be seen main difference here is between model with only control variables and model with control variables and key predictor - latter is slightly better than model with only control variables so it can be assumed that usage of key predictor, trust in legal system, is justified.


```{r results="asis", message=FALSE, warning=FALSE}
stargazer(anova(model, model1), type = "html", title="ANoVA", intercept.bottom = FALSE,
          single.row = TRUE, align=TRUE, ci=TRUE, ci.level=0.95, no.space=TRUE)
```

## Main model

```{r}
tab_model(model, title = "Main model", dv.labels = c("predictor & control"))
```

**Interpretation**:   
- *trust in the legal system* significantly leads to more planning attitude (more trust, more planning)!    
- *age* too have significant effect, but against my hypothesis ()   
- *gender*: have significant effect - females less plan for future than males.    
- *child(ren)*: not significant.    

# Assumptions

```{r}
plot_model(model, type = "diag")[[1]]
```

We don't have any multicollinearity!

```{r message=FALSE, warning=FALSE}
plot_model(model, type = "diag")[[2]]
```

And even residuals distributed normally and there are no major outliers!

```{r}
plot_model(model, type = "diag")[[3]]
```

Here again - residuals at the best shape and very nicely fits under the normal distribution line.

```{r message=FALSE, warning=FALSE}
plot_model(model, type = "diag")[[4]]
```

Again - homoscedasticity just not for my models :(

## Viz

```{r message=FALSE, warning=FALSE}
plot_summs(model, plot.distributions = TRUE, inner_ci_level = .9)
```

Plot shows all that already was described in Interpretation.


## Effects

```{r}
plot_model(model, type = "pred")[[1]]
```

**Trust in the legal system** has significant negative effect and with greater trust people plan more (or with greater trust people less 'take each day as it comes')

```{r}
plot_model(model, type = "pred")[[2]]
```

**Age** have significant effect and with more years being on Earth people tend to plan less and more 'take each day as it comes' - *some wisdom is hidden here*


```{r}
plot_model(model, type = "pred")[[3]]
```

And **gender** - females more 'take each day as it comes' than males! (that's interesting)


# Summary

I came to the conclusion that it is really interesting topic to research - what influences on peoples attitudes to 'planning for future' or to 'take each day as it comes'. I suppose that it is highly connected with perceived stability in the society besides self-efficacy and some minor factors.

