---
title: "Logistic Regression (DataCamp course)"
author: "dts"
date: "10 02 2020"
output: html_document
---

# Parallel slopes

**y ~ x + z**, where x is numeric and z is categoric.
```{r plotting parallel slopes}
mod <- lm(hwy ~ displ + factor(year), data = mpg)
summary(mod)
# get the coefficients
library(broom)
augment(mod)

# plotting by fitted values, not by observed
ggplot(augment(mod), aes(x = displ, y = hwy, color = factor.year.)) + 
  geom_point() +
  geom_line(aes(y = .fitted))
```

"After controlling for ..." - proper way to describe such model.


# Evaluating and extending parallel slopes model

Residual = real - fitted (values)

## Model fit

**Recall** (or *coefficient of determination*): R^2 = 1 - SSE/SST 
- *SSE* = the Sum of the Squared Errors (or residuals) 
- *SST* = the Total Sum of the Squares 

*Additional explanatory variables always increase recall (R^2)!*

**Recall adjusted**: R^2adj = 1 - SSE/SST * n-1/n-p-1
- *p* = number of explanatory variables

```{r fitted values}
# return a vector of fitted values
predict(mod)
# return a data.frame
augment(mod)
```

## Predicting
```{r predicting for new data}
new_obs <- data.frame(displ = 1.8, year = 2008)
predict(mod, newdata = new_obs)
augment(mod, newdata = new_obs)
```

## Interaction

Interaction term = 

```{r Interaction}
# only graphical representation
ggplot(data = mpg, aes(x = displ, y = hwy, color = factor(year))) +
  geom_point() + 
  geom_smooth(method = "lm", se = 0)

# creating a model
mod2 <- lm(hwy ~ displ + factor(year) + displ:factor(year), data = mpg)
summary(mod2)
# using fitted values from a model
ggplot(augment(mod2), aes(x = displ, y = hwy, color = factor.year.)) + 
  geom_point() +
  geom_line(aes(y = .fitted))
```

### Simpson's Paradox

Subgroups are important!


# Multiple Regression

```{r}
library(UsingR); data(babies)
babies <- subset(babies, age != 99 & gestation != 999)

# Adding z-axe as color
ggplot(babies, aes(x = gestation, y = age)) + 
  geom_point(aes(color = wt))

# Tiling the plane
library(dplyr); library(modelr)
grid <- babies %>%
  data_grid(
    gestation = seq_range(gestation, by = 1), 
    age = seq_range(age, by = 1)
  )
mod <- lm(wt ~ gestation + age, data = babies)
bwt_hats <- broom::augment(mod, newdata = grid)

ggplot(babies, aes(x = gestation, y = age)) + 
  geom_point(aes(color = wt)) + 
  geom_tile(data = bwt_hats, aes(fill = .fitted, alpha = 0.5)) + 
  scale_fill_continuous("wt", limits = range(babies$wt))

# 3D visualization
# library(plotly)
# plot_ly(data = babies, z = ~wt, x = ~gestation, y = ~age, opacity = 0.6) %>%
#   add_markers(text = ~case, marker = list(size = 2)) %>%
#   add_surface(x = ~x, y = ~y, z = ~plane, showscale = FALSE,
              # cmax = 1, surfacecolor = color1, colorscale = col1)
```

## Conditional interpretation of coefficients

Each B coefficient in model should be interpreted "**Outcome raises/decreases by N when X is raises/decreases by 1 while Y is constant (with mean of Y)**"

```{r adding third categorical predictor}
lm(wt ~ gestation + age + smoke, babies)
```

```{r}
# library(plotly)
# plot_ly(data = babies, z = ~wt, x = ~gestation, y = ~age, opacity = 0.6) %>%
#   add_markers(color = ~factor(smoke), text = ~"kek", marker = list(size = 2)) %>%
#   add_surface(x = ~x, y = ~y, z = ~plane0, showscale = FALSE, 
#               cmin = 0, cmax = 1, surfacecolor = "blue", colorscale = "blue") %>%
#   add_surface(x = ~x, y = ~y, z = ~plane1, showscale = FALSE, 
#               cmin = 0, cmax = 1, surfacecolor = "red", colorscale = "red")
```


# Logistic Regression

**geom_jitter** helps to visualize dots on graph with some distance between them on Y axis when point plotted only by X axes

```{r}
# ggplot(data = heartTr, aes(x = age, y = survived)) + 
#   geom_jitter(width = 0, height = 0.05, alpha = 0.5)
```

```{r}
# binned points and line
data_space <- ggplot(data = MedGPA_binned, aes(x = mean_GPA, y = acceptance_rate)) + geom_point() + geom_line()

# augmented model
MedGPA_plus <- mod %>%
  augment(type.predict = "response")

# logistic model on probability scale
data_space +
  geom_line(data = MedGPA_plus, aes(x = GPA, y = .fitted), color = "red")
```

## Converting
```{r}
# probability
heartTr_plus <- mod %>%
  augment(type.predict = "response") %>%
  mutate(y_hat = .fitted)

ggplot(heartTr_plus, aes(x = age, y = y_hat)) + 
  geom_point() + geom_line() + 
  scale_y_continuous("Probability of being alive", limits = c(0, 1))

# odds
heartTr_plus <- heartTr_plus %>%
  mutate(odds_hat = y_hat / (1 - y_hat))

ggplot(heartTr_plus, aes(x = age, y = odds_hat)) + 
  geom_point() + geom_line() + 
  scale_y_continuous("Odds of being alive")

# logit
heartTr_plus <- heartTr_plus %>%
  mutate(log_odds_hat = log(odds_hat))

ggplot(heartTr_plus, aes(x = age, y = log_odds_hat)) + 
  geom_point() + geom_line() + 
  scale_y_continuous("Log(odds) of being alive")

# odds ratios
exp(coef(mod))
```

```{r task on converting}
# compute odds for bins
MedGPA_binned <- MedGPA_binned %>%
  mutate(odds = acceptance_rate / (1 - acceptance_rate))

# plot binned odds
data_space <- ggplot(MedGPA_binned, aes(x = mean_GPA, y = odds)) + 
  geom_point() + geom_line()

# compute odds for observations
MedGPA_plus <- MedGPA_plus %>%
  mutate(odds_hat = .fitted / (1 - .fitted))

# logistic model on odds scale
data_space +
  geom_line(data = MedGPA_plus, aes(x = GPA, y = odds_hat), color = "red")


# compute log odds for bins
MedGPA_binned <- MedGPA_binned %>% mutate(log_odds = log(acceptance_rate / (1 - acceptance_rate)))

# plot binned log odds
data_space <- ggplot(data = MedGPA_binned, aes(x = mean_GPA, y = log_odds)) + geom_point() + geom_line()

# compute log odds for observations
MedGPA_plus <- MedGPA_plus %>% mutate(log_odds_hat = log(.fitted / (1 - .fitted)))

# logistic model on log odds scale
data_space +
  geom_line(data = MedGPA_plus, aes(x = GPA, y = log_odds_hat), color = "red")
```

## Converting coefficients
```{r}
# log-odds scale of fitted values
broom::augment(mod)
# probability scale of fitted values
broom::augment(mod, type.predict = "response")


# Out-of-sample predictions
cheney <- data.frame(age = 71, transplant = "treatment")
augment(mod, newdata = cheney, type.predict = "response")
```


```{r}
# Making binary predictions
mod_plus <- augment(mod, type.predict = "response") %>%
  mutate(alive_hat = round(.fitted))

mod_plus %>%
  select(is_alive, age, transplant, .fitted, alive_hat)
```


# case: Italian restaurants in NYC

```{r}
nyc <- read.csv(url("https://assets.datacamp.com/production/repositories/845/datasets/639a7a3f9020edb51bcbc4bfdb7b71cbd8b9a70e/nyc.csv"))
dplyr::glimpse(nyc)
pairs(nyc)
```

## Assumptions

- check for multicollinearity 

```{r}
library(plotly)
# draw 3D scatterplot
p <- plot_ly(data = nyc, z = ~Price, x = ~Food, y = ~Service, opacity = 0.6) %>%
  add_markers(color = ~factor(East)) 

# draw two planes
p %>%
  add_surface(x = ~x, y = ~y, z = ~plane0, showscale = FALSE) %>%
  add_surface(x = ~x, y = ~y, z = ~plane1, showscale = FALSE)
```


