---
title: "Cluster Analysis - DataCamp"
author: "dts"
date: "04 02 2020"
output: html_document
---

Steps:
1. No missing values & all on one scale
2. Choose basis for clustering
3. Clustering!
4. What those cluster mean?
(repeat from step 2 if need to)


# Distance vs. Similarity

Distance = 1 - Similarity

Dissimilarity can be counted with **Euclidean distance** (c^2 = a^2 + b^2):
```{r Euclidean distance}
# with 2 observations
(two_players = matrix(c(0,9,0,12), nrow = 2, dimnames = list(c('BLUE', 'RED'), c('X', 'Y'))))
dist(two_players, method = 'euclidean')
# with more than 2 observations
(three_players = matrix(c(0,9,-2,0,12,19), nrow = 3, dimnames = list(c('BLUE', 'RED', 'GREEN'), c('X', 'Y'))))
dist(three_players, method = 'euclidean')
```


# Standartization (scaling)

Xscaled = X - mean(X)/sd(X)

After standartization of all features they all will have *mean = 0 and sd = 1*
```{r scaling}
(heigth_weight = data.frame(Height = c(6,6,8), Weight = c(200,202,200)))
scale(heigth_weight)
```

# Distance for categorical data

**Jaccard index** - J(A,B) = intersection of A & B / union of A & B

Distance = 1 - Jaccard index

```{r Jaccard index}
(survey_a <- data.frame(wine = c(T,F,T), beer = c(T,T,F), whiskey = c(F,T,T), vodka = c(F,T,F)))
dist(survey_a, method = "binary")
```

More than two categories -> use **dummification** (recode all into logical)
```{r Dummification}
(survey_b <- data.frame(color = c('red','green','blue','blue'), sport = c('soccer', 'hockey', 'hockey', 'soccer')))
library(dummies)
(dummy_survey_b <- dummy.data.frame(survey_b))
dist(dummy_survey_b, method = "binary")
```
