---
title: "DA2020 - cluster analysis project (Tsimokha Dmitriy, 161)"
author: "Dmitriy Tsimokha"
date: "3/14/2020"
output:
  pdf_document: default
  html_document: default
---

Your task is to clusterise R users among the participants of the Stackoverflow 2019 survey for a software company targeting R users. "R users" are defined here those who indicated they worked with the R language. Using the variables of your choice, explore the possible clustering solutions in this group, and describe the differrences between the obtained clusters of R users.

Questionnaire: Stackoverflow 2019

Data:  [dataforproject2.csv](https://lms.hse.ru/view_file.php?action=download&file=2166667), attached here (filtered data from Stackoverflow 2019)
Deadline: March 10th, 11:50 PM. Later submissions have 50% off the grade.

The project should be submitted as an archive with **2 files: the Rmd script and the html output**. As an alternative, you can include in your script a link to the rpubs publication or another external repo. The report should not be too lengthy or overstuffed (up to 10 A4 pages long).

All projects will be graded according to the following criteria:

  1) data are described (name 1-3 important characteristics of your data) - 0.6 points
  2) variable choice is justified (you may want to use not all the variables) - 1 point
  3) the distance metric matches variable types - 1 point (if this is incorrect, interpretation will fail)
  4-5) you have tried a combination of k-means/PAM and agglomerative clustering - 2 points
  6) clusters have been visualized - 1 point
  7) decision on how many clusters to retain is based on tests/ dendrograms - 1 point
  8) comments on which clustering method delivers better results - 1 point
  9) the resulting clusters are described and labelled - 1.4 points
  10) the cluster solution is validated with any other methods (including but not limited to: PCA, regression, etc.) - 1 point

Original solutions are especially welcomed.

Grade descriptions:

  * 9-10 clusters are defined, described, and named the report is coherent and complete 
  * 8 clusters are defined, described, and named but 1-2 minor mistakes are allowed
  * 7 clusters are defined and described but a minor part is missing or the report contains several minor mistakes 
  * 6 clusters are defined and described but a minor part is missing or at least one serious mistake is made 
  * 5 clusters are defined but not described, the report is not coherent
  * 4 clusters are defined, analysis has flaws in logic, but it is structured (distances are calculated, clusters are identified, and results are summarised) 
  * 3 clusters are not identified, no summary provided
  
```{r}
library(data.table)
df <- fread("dataforproject2.csv", stringsAsFactors = T)
str(df)
```
  
To clusterize R user meaningfully I take following variables:

  * Age
  * Extraversion
  * OpSys
  * 
  
```{r}
sapply(df[,c("Age", "Extraversion", "OpSys")], summary)
```
  
  